Carbonate chemistry at Ny-Ålesund

### 2022-08-08 awipev-CO2.Rmd
misc

```{r read and prepare discrete AT+CT, echo=FALSE, warning=FALSE, message=FALSE}
d <-
  read_excel(
    paste0(
      path,
      "fb_data/Discrete_analyses_AT_CT/Discrete_sampling_AWIPEV.xlsx"
    ),
    skip = 0,
    col_names = TRUE,
    na = "NA"
  ) %>%
  dplyr::mutate(
    datetime = lubridate::dmy_hm(sampling_date),
    measure_date = lubridate::dmy(measure_date, tz = "UTC"),
    date = lubridate::as_date(datetime)
  ) %>%
  dplyr::rename(
    ph_s_sf = pH_s_seafet,
    ph_s_dur = pH_s_durafet,
    qflag_ph_s_sf = qflag_pH_s_seafet,
    qflag_ph_s_dur = qflag_pH_s_durafet
  ) %>%
  dplyr::select(datetime, date, everything())
# Add the flag location: some samples have been collected at the pier. Notes them here.
# Flag 1 = collected on the pIer by Niskin and flag 0 = collected in the FB (normal collect)
 # And keep data with flag 0 for FB location
d <- d %>%
  dplyr::mutate(location_flag = ifelse(datetime == as.POSIXct("2016-04-21 09:25:00") | 
                                         datetime == as.POSIXct("2016-02-10 14:00:00") | 
                                         datetime == as.POSIXct("2016-02-17 09:20:00") |
                                         datetime == as.POSIXct("2016-02-24 14:20:00") |
                                         datetime == as.POSIXct("2016-02-04 09:30:00") |
                                         datetime == as.POSIXct("2016-01-27 15:00:00") |
                                         datetime == as.POSIXct("2016-01-13 13:45:00") |
                                         datetime == as.POSIXct("2016-01-06 14:10:00") |
                                         datetime == as.POSIXct("2015-12-24 10:45:00") |
                                         datetime == as.POSIXct("2015-12-30 13:30:00") |
                                         datetime == as.POSIXct("2018-01-05 15:20:00") |
                                         datetime == as.POSIXct("2018-02-02 14:35:00") |
                                         datetime == as.POSIXct("2018-03-02 14:10:00") |
                                         datetime == as.POSIXct("2019-05-31 11:45:00"), 
                                       1, 0))
d <- d %>%
  dplyr::filter(location_flag == 0) %>% 
# below we change the time of 2 discrete samplings during period 1 of seafet deployment
# This is to enable calibration when the REMOS was at 9 m depth
  dplyr::mutate(datetime = case_when(datetime == as.POSIXct("2017-08-24 13:30:00") ~ as.POSIXct("2017-08-24 19:00:00"),
                                     datetime == as.POSIXct("2017-11-28 09:15:00") ~ as.POSIXct("2017-11-28 13:00:00"),
                   TRUE ~ datetime))

@@

```{r measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
reg <- ggreg(
  xdata = d_all$pco2_fb_insi_temp,
  ydata = d_all$pco2_calc_fb_insi_temp,
  type = "MA",
  plot = TRUE,
  title = "pCO2 normalized at in situ temperature",
  #subtitle = "",
  xlab = latex2exp::TeX("Measured $pCO_{2}\\phantom{x}(\\mu atm)$"),
  ylab = latex2exp::TeX("Calculated $pCO_{2}\\phantom{x}(\\mu atm)$"),
  proplot = TRUE,
  #ylim(c(30.5,35.5)) + xlim(c(28,35.5)),
  coord_fixed(ratio = 1)
)
print(reg$plot)
ggsave(filename = "figures/pco2_calc_meas.png", plot = reg$plot)
```

@@

```{r plot measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
print(reg$plot)
rm(reg)
```

@@

pH is measured continuously by an autonomous seaFET pH sensor.
Measurement periodicity was set to 1 hour from 2018-02-02 to 2018-04-17
it was then measured once every minute. pH calibration is based on
monthly, when possible, discrete samples collected at 11 m that were
analysed with spectrophotometric method in Villefranche by Samir
Alliouane. Measurments were performed several months after sampling.
When sampling was not possible in the vicinity of the sensor for
practical reasons (in winter), sampling was done in the harbour and is
not taken in consideration for calibration for spatial reasons. *Note
that the pH sensor is mounted on a profiling structure. Here, in
principle, only seaFET data recorded with a depth \> 8 m should be taken
into consideration.*

@@

# To make sf calibration against pH spectro, we need: sal - T insi - voltINT/EXT . If we missed one of these parameters, calibration will not run
# Here we try to fill the gap when it is possible: make a mean with 4 previous/next hours available in the time serie. 
# If the pH spectro is too far from a needed value, we decided to skip it for the calibration.

@@

#interpolation of salinity during gaps
# # below is old code. now we use sal_insitu_ctd_mix
# salinterp <-
#   approx(
#     d_all$datetime,
#     d_all$sal_insitu_ctd,
#     xout = d_all$datetime,
#     method = "linear",
#     rule = 2
#   )
# names(salinterp) <- c("datetime", "sal_insitu_ctd_interp")
# salinterp <- as.data.frame(salinterp)
# d_all$sal_insitu_ctd_interp <- salinterp$sal_insitu_ctd_interp
# 
# d_all <- d_all %>%
#   dplyr::mutate(
#     sal_insitu_ctd_interp =  ifelse(
#       datetime > as.POSIXct("2018-04-16 07:00:00") &
#         datetime < as.POSIXct("2018-04-24 22:00:00"),
#       sal_insitu_ctd_interp,
#       sal_insitu_ctd
#     )
#   )

# above is old code. Now we use sal_insitu_ctd_mix
salinterp <-
  approx(
    d_all$datetime,
    d_all$sal_insitu_ctd_mix,
    xout = d_all$datetime,
    method = "linear",
    rule = 2
  )
names(salinterp) <- c("datetime", "sal_insitu_ctd_mix_interp")
salinterp <- as.data.frame(salinterp)
d_all$sal_insitu_ctd_interp <- salinterp$sal_insitu_ctd_mix_interp

@@

# make the mean of E0 by period of deployement
calib_spectro_all <- group_by(calib_spectro_all, deployment_sf) %>%
  dplyr::summarise(E0int25 = mean(E0int25, na.rm = TRUE),
            E0ext25 = mean(E0ext25, na.rm = TRUE) 
            ) %>% 
  dplyr::filter(!is.na(deployment_sf))
names(calib_spectro_all) <- c("deployment_sf", "E0int25_sf_all", "E0ext25_sf_all")

@@

Final calibrated seaFET pHT data (small green dots) are shown here after
calibration with spectro measurements (red dots) according to each
period of deployement, defined between 2 spectro measurements (shading
grey).

@@

pH INT - pH EXT was plotted here. From now on, one uses seaFET pH from the internal electrode calibrated
with spec pH and seaFET data when the REMOS was below 8 m (pHint_tot_sf_10m). This unique pH variable is called pHint_tot_sf. All other pH variables are cleaned up. We also calculate pHint_tot_sf. at 4°C: pHint_tot_sf_temp_4.

>>>>

```{r read and prepare discrete AT+CT, echo=FALSE, warning=FALSE, message=FALSE}
d <-
  read_excel(
    paste0(
      path,
      "fb_data/Discrete_analyses_AT_CT/Discrete_sampling_AWIPEV.xlsx"
    ),
    skip = 0,
    col_names = TRUE,
    na = "NA"
  ) %>%
  dplyr::mutate(
    datetime = lubridate::dmy_hm(sampling_date),
    measure_date = lubridate::dmy(measure_date, tz = "UTC"),
    date = lubridate::as_date(datetime)
  ) %>%
  dplyr::rename(
    ph_s_sf = pH_s_seafet,
    ph_s_dur = pH_s_durafet,
    qflag_ph_s_sf = qflag_pH_s_seafet,
    qflag_ph_s_dur = qflag_pH_s_durafet
  ) %>%
  dplyr::select(datetime, date, everything())
# Add the flag location: some samples have been collected at the pier. Notes them here.
# Flag 1 = collected on the pIer by Niskin and flag 0 = collected in the FB (normal collect)
 # And keep data with flag 0 for FB location
d <- d %>%
  dplyr::mutate(
    location_flag = ifelse(
      datetime == as.POSIXct("2016-04-21 09:25:00") |
        datetime == as.POSIXct("2016-02-10 14:00:00") |
        datetime == as.POSIXct("2016-02-17 09:20:00") |
        datetime == as.POSIXct("2016-02-24 14:20:00") |
        datetime == as.POSIXct("2016-02-04 09:30:00") |
        datetime == as.POSIXct("2016-01-27 15:00:00") |
        datetime == as.POSIXct("2016-01-13 13:45:00") |
        datetime == as.POSIXct("2016-01-06 14:10:00") |
        datetime == as.POSIXct("2015-12-24 10:45:00") |
        datetime == as.POSIXct("2015-12-30 13:30:00") |
        datetime == as.POSIXct("2018-01-05 15:20:00") |
        datetime == as.POSIXct("2018-02-02 14:35:00") |
        datetime == as.POSIXct("2018-03-02 14:10:00") |
        datetime == as.POSIXct("2019-05-31 11:45:00"),
      1,
      0
    )
  )
d <- d %>%
  dplyr::filter(location_flag == 0) %>% 
# below we change the time of 2 discrete samplings during period 1 of seafet deployment
# This is to enable calibration when the REMOS was at 9 m depth
  dplyr::mutate(datetime = case_when(datetime == as.POSIXct("2017-08-24 13:30:00") ~ as.POSIXct("2017-08-24 19:00:00"),
                                     datetime == as.POSIXct("2017-11-28 09:15:00") ~ as.POSIXct("2017-11-28 13:00:00"),
                   TRUE ~ datetime))

@@

```{r measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
reg <- ggreg(
  xdata = d_all$pco2_fb_insi_temp,
  ydata = d_all$pco2_calc_fb_insi_temp,
  type = "MA",
  plot = TRUE,
  title = "pCO2 normalized at in situ temperature",
  xlab = latex2exp::TeX("Measured $pCO_{2}\\phantom{x}(\\mu atm)$"),
  ylab = latex2exp::TeX("Calculated $pCO_{2}\\phantom{x}(\\mu atm)$"),
  proplot = TRUE,
  coord_fixed(ratio = 1)
)
print(reg$plot)
pco2_calc_meas <- reg$plot
ggsave(filename = "figures/pco2_calc_meas.png", plot = reg$plot)
```

@@

# ```{r plot measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
# print(reg$plot)
# rm(reg)
# ```

@@

pH is measured continuously by an autonomous seaFET pH sensor.
Measurement periodicity was set to 1 hour from 2018-02-02 to 2018-04-17
it was then measured once every minute. pH calibration is based on
monthly, when possible, discrete samples collected at 11 m that were
analysed with spectrophotometric method in Villefranche by Samir
Alliouane. Measurements were performed several months after sampling.
When sampling was not possible in the vicinity of the sensor for
practical reasons (in winter), sampling was done in the harbour and is
not taken in consideration for calibration for spatial reasons. *Note
that the pH sensor is mounted on a profiling structure. Here, in
principle, only seaFET data recorded with a depth \> 8 m should be taken
into consideration.*

@@

# To make sf calibration against pH spectro, we need: sal - T insi - voltINT/EXT . If we missed one of these parameters, calibration will not run
# Here we try to fill the gap when it is possible: make a mean with 4 previous/next hours available in the time series. 
# If the pH spectro is too far from a needed value, we decided to skip it for the calibration.

@@

salinterp <-
  approx(
    d_all$datetime,
    d_all$sal_insitu_ctd_mix,
    xout = d_all$datetime,
    method = "linear",
    rule = 2
  )
names(salinterp) <- c("datetime", "sal_insitu_ctd_mix_interp")
salinterp <- as.data.frame(salinterp)
d_all$sal_insitu_ctd_interp <- salinterp$sal_insitu_ctd_mix_interp

@@

# make the mean of E0 by period of deployment
calib_spectro_all <- group_by(calib_spectro_all, deployment_sf) %>%
  dplyr::summarise(E0int25 = mean(E0int25, na.rm = TRUE),
            E0ext25 = mean(E0ext25, na.rm = TRUE) 
            ) %>% 
  dplyr::filter(!is.na(deployment_sf))
names(calib_spectro_all) <- c("deployment_sf", "E0int25_sf_all", "E0ext25_sf_all")

@@

Final calibrated seaFET pHT data (small green dots) are shown here after
calibration with spectro measurements (red dots) according to each
period of deployment, defined between 2 spectro measurements (shading
grey).

@@

pH INT - pH EXT was plotted here. From now on, one uses seaFET pH from the internal electrode calibrated with spec pH and seaFET data when the REMOS was below 8 m (pHint_tot_sf_10m). This unique pH variable is called pHint_tot_sf. All other pH variables are cleaned up. We also calculate pHint_tot_sf at 4°C: pHint_tot_sf_temp_4.
